### 2021-07-31 paper/localpsf.tex
edits based on comments by J

To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\varphi_{x_i}(z)$ for a large number of points $z$. Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\eta_b(z+x_i)$. This is not ideal. In computations, $\eta^b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh. Evaluating $\eta_b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell. Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.

@@

Evaluating the kernel for our product-convolution approximation, $\AkerPc(y,x)$, requires evaluating $\varphi_{x_i}(y-x)$. But $\varphi_{x_i}(y-x)$ is undefined  when $y - x + x_i$ is outside $\Omega$, which may occur even if $x \in \Omega$ and $y \in \Omega$. The default approach to deal with this problem is to extend $\varphi_{x_i}$ by zero wherever it is undefined. But this leads to substantial errors, known as boundary artifacts, which are especially significant when the sample point $x_i$ is near the boundary of the domain. In this case, the impulse response is artificially chopped off by the domain boundary. See Figure FIG for an illustration, and see \cite{ADAPTIVECONVMYPAPER} for more detailed discussion of boundary artifacts.

@@

In detail, let $x_{n_1}, x_{n_2}, \dots, x_{n_k}$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself. That is, $x_{n_1}=x_i$ (the point itself), $x_{n_2}$ is the nearest sample point to $x_i$ other than $x_i$ itself, $x_{n_3}$ is the second nearest sample point to $x_i$ other than $x_i$ itself, and so on. Also, recall that in computations that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids. To extend $\varphi_{x_i}$, first we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels $\varphi_{x_{n_1}}, \varphi_{x_{n_2}}, \dots, \varphi_{x_{n_k}}$. We initialize an extended convolution kernel, $\varphi_{x_i}^E$, to zero on this expanded grid. Then we interpolate the functions $\varphi_{x_{n_j}}$ onto this expanded grid one at a time in reverse order. That is, first we interpolate $\varphi_{x_{n_k}}$ onto the grid, then $\varphi_{x_{n_{k-1}}}$, and so on until we reach the original kernel $\varphi_{x_{n_1}} = \varphi_{x_i}$. At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{x_{n_j}}$ is undefined (``NaN''). At gridpoints where the kernel being interpolated is undefined, we leave $\varphi_{x_i}^E$ unchanged. In this way, we fill in undefined values of $\varphi_{x_i}$ with values from successively further and further away kernels. This process is illustrated in Figure FIG. To avoid notational clutter, and since we will henceforth only work with the extended kernel, $\varphi_{x_i}^E$, in the remainder of this paper the symbol $\varphi_{x_i}$ refers to $\varphi_{x_i}^E$.

>>>>

To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\varphi_{x_i}(z)$ for a large number of points $z$. Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\eta_b(z+x_i)$. In computations, $\eta^b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh. Evaluating $\eta_b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell. Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.

@@

Evaluating the kernel for our product-convolution approximation, $\AkerPc(y,x)$, requires evaluating $\varphi_{x_i}(y-x)$. But $\varphi_{x_i}(y-x)$ is undefined  when $y - x + x_i$ is outside $\Omega$, which may occur even if $x \in \Omega$ and $y \in \Omega$. The default approach to deal with this problem is to extend $\varphi_{x_i}$ by zero wherever it is undefined. But this leads to substantial errors, known as boundary artifacts, which are especially significant when the sample point $x_i$ is near the boundary of the domain. In this case, the impulse response is artificially chopped off by the domain boundary. See Figure FIG for an illustration.

@@

In detail, let $p_1, p_2, \dots, p_k$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself. That is, $p_1=x_i$, $p_2$ is the nearest neighbor to $x_i$, $p_3$ is the second nearest neighbor to $x_i$, and so on. Also, recall that in computations that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids. To extend $\varphi_{x_i}$, first we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\varphi_{p_1}, \varphi_{p_2}, \dots, \varphi_{p_k}$. We initialize an extended convolution kernel, $\varphi_{x_i}^E$, to zero on this expanded grid. Then we interpolate the functions $\varphi_{p_j}$ onto this expanded grid one at a time in reverse order. That is, first we interpolate $\varphi_{p_k}$ onto the grid, then $\varphi_{p_{k-1}}$, and so on until we reach the original kernel $\varphi_{p_1} = \varphi_{x_i}$. At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{p_j}$ is undefined (``NaN''). At gridpoints where the kernel being interpolated is undefined, we leave $\varphi_{x_i}^E$ unchanged. In this way, we fill in undefined values of $\varphi_{x_i}$ with values from successively further and further away kernels. This process is illustrated in Figure FIG. To avoid notational clutter, and since we will henceforth only work with the extended kernel, $\varphi_{x_i}^E$, in the remainder of this paper the symbol $\varphi_{x_i}$ refers to $\varphi_{x_i}^E$.
